Placeholder Tex Doc

### 2019-01-09 tex_files/placeholder.tex
...

\section{Traditional Solutions}
A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques. The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees. %TODO insert figure
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
	\label{fig:trees}
\end{figure}
While on the surface these appear simple the task of obtaining an exact matching is, in practise, the limiting factor in this design. Even if the clock distribution system is geometrically symmetrical production mismatches in either active or passive components will lead to a skew that varies from part to part. In order to minimise the impact of production tolerances the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible. However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.
Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses however this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.

A mesh clock distribution network is an alternate design where the clock is delivered using a Cartesian grid of distribution lines and variation in skew is inversely proportional to the density of the grid. According to Abdelhadi \textit{et al} (2010) clock meshes ``\textit{achieve low and deterministic skew, low skew variations, and low jitter}'', all desirable characteristics for a clock distribution system. However they dissipate more power due to extra capacitive loading due to vast number of lines required to form the grid and similarly to tree distribution networks suffer from potential mismatch in production and alleviation through increasing dimensions will again lead to higher power consumption \cite{abdelhadi2010timing}. 
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}%TODO insert figure
	\label{fig:mesh}
\end{figure}

@@

\begin{figure}[h] %TODO top
	\centering
	\includegraphics[scale=0.35]{../shan_bb_pd}
	\caption{Bang-bang phase/frequency detector architecture. \cite{shan2014phd}.}
	\label{fig:shan_bb_pd}
\end{figure}
\subsection{Digital Loop Filter}
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{../shan_bb_pd} %TODO this diagram
	\caption{Basic PI Controller Architecture.}
	\label{fig:my_simple_pi}
\end{figure}
The loop filter in an ADPLL can be implemented as a PI controller as it is just a low pass filter, such as that in Figure \ref{fig:my_simple_pi}. In the case of an ADPLL node in a network the input of this filter is a weighted sum of two to four phase errors, each corresponding to a neighbouring node. In a digital system the proportional section is implemented by a simple multiplier whereas the integral path is a combination of a multiplier and a delayed summation stage. The value of these gains determine the response and stability of the oscillator system. The transfer function of such a circuit is given by:
\begin{equation*}
	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}~~~\cite{shan2014phd}
\end{equation*}
It has been found by Koskin \textit{et al} (2018) that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}. In the same study a range of values were found which would produce low jitter operation of the network. As these value are all less than one the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.

@@

\section{ADPLL}
Combining these blocks, in this case using the Phase Accumulator style oscillator, gives the ADPLL shown in Figure \ref{fig:my_adpll} on which I performed my initial testing and simulations.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{../rtl}
	\caption{Phase Accumulator ADPLL RTL diagram.}
	\label{fig:my_adpll}
\end{figure}

\chapter{Test Results}

\chapter{Future Work}
So far all work has been done using a single phase lock loop in order to have a suitable basis for the network and going forward this PLL will be used in that role. The first stage of network implementation will interlink two PLLs, with one of them also receiving input from the external reference. This test will serve as both validation of the phase coupling ability of the network as well as a test of the error summation block and the ability to lock from startup. For the sake of more realistic performance the inverter chain ADPLL will be used ahead of the Phase Accumulator based design.\\
A shortcoming of the current FPGA evaluation board is the inability to input or extract signals in the multiple MHz range without significant distortion, and this will be resolved with the arrival of an FPGA more suited to extracting and importing higher frequency signals. With an ability to extract higher frequency signals for measurement and analysis purposes the target frequency of 5 MHz can potentially be increased into the tens of MHz range.\\
With simple coupling implemented the logical next stage is the implementation of a small network, possibly 2x2, before scaling up to a size suitable to the FPGA in use. At this point analysis of measurement results can be undertaken and any refinements to the network can be made. Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.

>>>>

\section{Traditional Solutions}
A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques. The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees. %TODO insert figure
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.3]{../trees}
	\caption{H and Branch Tree Clock Distribution Systems.}%TODO insert figure
	\label{fig:trees}
\end{figure}
While on the surface these appear simple the task of obtaining an exact matching is, in practise, the limiting factor in this design. Even if the clock distribution system is geometrically symmetrical production mismatches in either active or passive components will lead to a skew that varies from part to part. In order to minimise the impact of production tolerances the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible. However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.
Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses however this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.

A mesh clock distribution network is an alternate design where the clock is delivered using a Cartesian grid of distribution lines and variation in skew is inversely proportional to the density of the grid. According to Abdelhadi \textit{et al} (2010) clock meshes ``\textit{achieve low and deterministic skew, low skew variations, and low jitter}'', all desirable characteristics for a clock distribution system. However they dissipate more power due to extra capacitive loading due to vast number of lines required to form the grid and similarly to tree distribution networks suffer from potential mismatch in production and alleviation through increasing dimensions will again lead to higher power consumption \cite{abdelhadi2010timing}. 
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.6]{eldar_mesh}
	\caption{Mesh Clock Distribution System \cite{zianbetov2013phd}.}%TODO insert figure
	\label{fig:mesh}
\end{figure}

@@

\begin{figure}[h] %TODO top
	\centering
	\includegraphics[scale=0.35]{../shan_bb_pd}
	\caption{Bang-bang phase/frequency detector architecture. \cite{shan2014phd}.}
	\label{fig:shan_bb_pd}
\end{figure}
\subsection{Digital Loop Filter}
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.45]{../pi_simple} %TODO this diagram
	\caption{Basic PI Controller Architecture.}
	\label{fig:my_simple_pi}
\end{figure}
The loop filter in an ADPLL can be implemented as a PI controller as it is just a low pass filter, such as that in Figure \ref{fig:my_simple_pi}. In the case of an ADPLL node in a network the input of this filter is a weighted sum of two to four phase errors, each corresponding to a neighbouring node. In a digital system the proportional section is implemented by a simple multiplier whereas the integral path is a combination of a multiplier and a delayed summation stage. The value of these gains determine the response and stability of the oscillator system. The transfer function of such a circuit is given by:
\begin{equation*}
	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}~~~\cite{shan2014phd}
\end{equation*}
It has been found by Koskin \textit{et al} (2018) that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}. In the same study a range of values were found which would produce low jitter operation of the network. As these value are all less than one the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.

@@

\section{ADPLL}
Combining these blocks, in this case using the Phase Accumulator style oscillator, gives the ADPLL shown in Figure \ref{fig:my_adpll} on which I performed my initial testing and simulations. For ease of testing the PLL replicated the reference exactly, as this would facilitate easier testing of the lock and capture ranges of the design.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{../rtl}
	\caption{Phase Accumulator ADPLL RTL diagram.}
	\label{fig:my_adpll}
\end{figure}\\
Once this testing had been completed with the phase accumulator based ADPLL testing progressed to the design implementing an inverter chain and scaling from the reference using an eight times divider. This necessitated a change in the bit width of the control signal to match the 5 bit wide signal desired, thus leading to the design in Figure \ref{fig:my_ring_adpll}.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{../ro_rtl}
	\caption{Ring Oscillator ADPLL RTL diagram.}
	\label{fig:my_ring_adpll}
\end{figure}\\
 The Loop Filter is clocked at the generated output frequency as this ensures it will be synchronised with the oscillator. This is important when using the inverter chain oscillator in order to avoid metastability as the generated output clock it may not be synchronised with the high frequency clock of the FPGA.

\chapter{Test Results}
\chapter{Test Results}

\chapter{Future Work}

So far all work has been done using a single phase lock loop in order to have a suitable basis for the network and going forward this PLL will be used in that role. The first stage of network implementation will interlink two PLLs, with one of them also receiving input from the external reference. This test will serve as both validation of the phase coupling ability of the network as well as a test of the error summation block and the ability to lock from startup. For the sake of more realistic performance the inverter chain ADPLL will be used ahead of the Phase Accumulator based design.\\
A shortcoming of the current FPGA evaluation board is the inability to input or extract signals in the multiple MHz range without significant distortion, and this will be resolved with the arrival of an FPGA more suited to extracting and importing higher frequency signals. With an ability to extract higher frequency signals for measurement and analysis purposes the target frequency of 5 MHz can potentially be increased into the tens of MHz range.\\
With simple coupling implemented the logical next stage is the implementation of a small network, possibly 2x2, before scaling up to a size suitable to the FPGA in use. At this point analysis of measurement results can be undertaken and any refinements to the network can be made. Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.\\
As with the single ADPLL testing the important characteristics to look out for will be the lock and capture ranges of the network which should mirror that of the ADPLLs used in it's construction. Similarly the jitter of a number of clocks across the network will be examined in order, and any skew between the PLL outputs investigated. Also of interest will be the initial locking behaviour, in particular how quickly the network can be brought up from ``cold''.
